Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2019-03-22 article.tex
Add notes
Feedback Frans

Transmembrane helices (TMHs) are an overlooked source of 
major histocompatibility complex (MHC) class I epitopes.
It is unknown why this is the case and 
whether MHC-II also is an overlooked source of epitopes.
This study investigates if there is an evolutionary conservation 
of a pathogen's TMHs and if this is matches 
with MHC-II's over/equally/under-presenting TMHs.
We find that ...
\end{abstract}

@@

For MHC-1, we know that epitopes derived from transmembrane helices (TMHs) are 
over-presented by all human leukocyte antigen (HLA)-A and most HLA-B super 
types [\cite{bianchi2017transmembrane}].

A hypothesis offered by that same paper is that the presentation of 
especially TMHs may have an evolutionary advantage for the (human) host.
The argumentation given is that the presentation of parts of TMHs
is likelier to detect pathogenic proteins, as TMHs are already conserved
due to the functional requirement of being able to span a membrane.
Additionally, there are fewer escape mutations possible for the pathogen,
as relatively less of these will result in a functional TMH.

However, that argumentation can also go the other way: as there are less
polypeptide fragments from (the less amount of) possible TMHs, 
it is likelier that host and pathogen
share the same protein sequence, 
resulting the pathogen avoiding detection.

Within the immune response, MHC-I is unique in [doing something].
The other most important participant in the immune response in MHC-II,
which is more involved in detecting bacterial pathogens, 
e.g. tuberculosis, [and others].

From the earlier research done on MHC-I, it naturally follows to check
if MHC-II are likelier to present TMHs than expected by chance.

Extending the previous study on MHC-I, we also measure the evolutionary
conservation of TMH epitopes in Mycoplasma pathogens.

@@

>>>>

Transmembrane helices (TMHs) are an overrepresented source of epitopes on major histocompatibility complex (MHC) class I due to a preference for hydrophobic for the majority of HLA-I haplotypes. It is unknown why there is such a preference and whether  this is the case for MHC-II haplotypes. This study investigates if there is an evolutionary conservation of a pathogen’s TMHs and if this correlates with the affinity for binding to MHC-II.

TMHs more/less/equal as expected by chance. We ﬁnd that ...

\end{abstract}

@@

For MHC-1, it was found that epitopes derived from transmembrane helices (TMHs)
are over-presented by all human leukocyte antigen (HLA)-A and most HLA-B super 
types [\cite{bianchi2017transmembrane}]. One explanation is that the 
presentation of especially TMHs may have an evolutionary advantage for 
the (human) host. The argumentation for this is that the TMHs are more 
conserved due to the functional requirement of being able to span a lipid 
bilayer. Additionally, there are limited escape mutations possible for the 
pathogen, as many will result in a non-functional TMH. 
However, that argumentation can also be countered as there are limited unique 
polypeptide fragments and TMHs are a conserved throughout all species, it is 
likelier that host and pathogen share the same unique epitope sequences, 
resulting in the pathogen avoiding detection due to the negative T-cell 
selection process in the thymus.

If presentation of TMH’s on MHC-I would bring an evolutionary advantage 
in the recognition of pathogens by the immune, 
it would follow that this is equally important for MHC-II. 
MHC-II, which is necessary for the activation of B-cells after the detection 
of bacterial pathogens, e.g.  tuberculosis, by dendritic cells or [...]. 
From the earlier research done on MHC-I, it naturally follows to check if 
also MHC-II is more likely to present TMHs than expected by chance. 
Extending the previous study on MHC-I, we also measure the evolutionary 
conservation of TMH epitopes in Mycoplasma pathogens to adress if 
presentation of TMH derived epitopes has an evolutionary advantage.

@@

\richel{
  Note that most bacteria are opportunistic pathogens.
  Note that most bacteria are generalists.
  Note that most bacteria have different cell membranes (and walls), that
  may have different functional constraints than a human cell membrane
}
